---
title: "An introduction to R Markdown"
author: "Fábio M. Corrêa <br/> **Email: <f.correa@ru.ac.za>"
site: bookdown::bookdown_site
output: 
   bookdown::gitbook:
      toc_depth: 4
documentclass: book
colorlinks: yes
monofont: "Source Code Pro"
---

```{r, echo=FALSE}
# Defining the same size for all figures
knitr::opts_chunk$set(echo=FALSE, fig.width=10, fig.height=8)
```


# Last update {-}
`r format(Sys.Date(), "%d-%m-%Y")`

 ***The text is under development and updates are constant***



# R Markdown Basics

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple
formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_
provides the flexibility of _Markdown_ with the implementation of **R** input
and output.  For more details on using _R Markdown_ see <https://bookdown.org/yihui/rmarkdown/>. 

Be careful with your spacing in _Markdown_ documents.  While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.  As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.  In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).

## Lists

It's easy to create a list.  It can be unordered like

* Item 1
* Item 2

or it can be ordered like

1. Item 1
4. Item 2

Notice that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.

To create a sublist, just indent the values a bit (at least four spaces or a tab).  (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
    - Item 3a
    - Item 3b

## Line breaks

Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:* 

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.

This should be a new paragraph.

## Font syntax


*italics* or _italics_

```{r, eval=FALSE, echo=TRUE}
*italics* or _italics_
```


```{r, eval=FALSE, echo=TRUE}
**bold** or __bold__
```

superscript^2^

```{r, eval=FALSE, echo=TRUE}
superscript^2^
```

subscript~2~

```{r, eval=FALSE, echo=TRUE}
superscript~2~
```


~~strikethrough~~

```{r, eval=FALSE, echo=TRUE}
~~strikethrough~~
```

## R chunks

When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document. You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):

The basic R chunk is:

````markdown
`r ''````{r cars}
summary(cars)
```
````

The result is:

```{r cars}
summary(cars)
```


### Chunk options

You can define chunk options globally or individually. A global definition can
be inserted into your file by calling the `knitr::opts_chunk$set` function. A global
definition can be modified by individual definitions.

#### Global options

````markdown
`r ''````{r}
knitr::opts_chunk$set(options)
```
````

#### Individual options

````markdown
`r ''````{r, options}
R code
```
````

#### Some chunk options


* Text output

	**echo**:(TRUE or FALSE) Whether to display the source code in the output
	document.

	**collapse**: (TRUE or FALSE) Collapse all the source and output blocks created by the chunk into a single block.

	**warning**: (TRUE or FALSE) Whether to preserve warnings (produced by warning())
	in the output.

	**error**: (TRUE or FALSE) Whether to preserve errors (from stop()).

	**eval**: (TRUE or FALSE) Defines whether R code should be run or not.


For a detailed list of options, see:

[link] <https://yihui.org/knitr/options/>

[link] <https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>



## Inline code

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:

```{r, echo=TRUE, eval=FALSE}
> The `cos` of $2 \pi$ is `r cos(2*pi)`. 
```


> The `cos` of $2 \pi$ is `r cos(2*pi)`. 

Another example would be the direct calculation of the standard deviation:

```{r, echo=TRUE, eval=FALSE}
> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.
```

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.


One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:


```{r, echo=TRUE, eval=FALSE}
> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`
```

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be indented.


# Math expressions

Mathematical expressions are written using **LaTeX** code.

Equations must be written between `$ $` or `$$ $$`. For inline equations you
must use `$<equation>$`.


## Inline equations


My first inline equation: $x = \frac{-b \pm \sqrt{b^{2} - 4ac}}{2a}$.


````markdown
My first inline equation: $x = \frac{-b \pm \sqrt{b^{2} - 4ac}}{2a}$.
````


## Math environments

### Equation


Equation \@ref(eq:ncrep)

\begin{equation}\label{NCntheta}
	\theta^{*}|L^{*}, \delta^{2},Y,\sigma^{2}_{u} \sim N\left(B^{-1}W'L^{*},\delta^{2}B^{-1} \right)
	(\#eq:ncrep)
\end{equation}


The equation reference is defined by `\@ref(eq:name) ` which must defined in
equation as `(\#eq:name)`


````markdown
Equation \@ref(eq:ncrep)

\begin{equation}
	\theta^{*}|L^{*}, \delta^{2},Y,\sigma^{2}_{u} \sim N\left(B^{-1}W'L^{*},\delta^{2}B^{-1} \right)
	(\#eq:ncrep)
\end{equation}
````


### Matrix


Matrix \@ref(eq:matex)

\begin{equation}
X_{m,n} = 
\begin{pmatrix}
  x_{1,1} & x_{1,2} & \cdots & x_{1,n} \\
  x_{2,1} & x_{2,2} & \cdots & x_{2,n} \\
  \vdots  & \vdots  & \ddots & \vdots  \\
  x_{m,1} & x_{m,2} & \cdots & x_{m,n} 
\end{pmatrix}
(\#eq:matex)
\end{equation}


````markdown

Matrix \@ref(eq:matex)


\begin{equation}
X_{m,n} = 
\begin{pmatrix}
  x_{1,1} & x_{1,2} & \cdots & x_{1,n} \\
  x_{2,1} & x_{2,2} & \cdots & x_{2,n} \\
  \vdots  & \vdots  & \ddots & \vdots  \\
  x_{m,1} & x_{m,2} & \cdots & x_{m,n} 
\end{pmatrix}
(\#eq:matex)
\end{equation}
````

An extensive list of examples and mathematical symbols can be viewed in the references.

## References

* <https://en.wikibooks.org/wiki/LaTeX/Mathematics>

* <https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html>



